The Etiology of Cybersecurity

### 2020-09-01 reports/paper_0/main.tex
anonymize

@@

\author{\IEEEauthorblockN{Marco Rocchetto}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: marco@v-research.it}
\and
\IEEEauthorblockN{Francesco Beltramini}
\IEEEauthorblockA{V-Research\\
Verona, Italy\\
Email: francesco@v-research.it}
}

@@

There is no agreement between the research communities (e.g.
Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
However, the same ideas revolved around for thousands of years\footnote{
It is interesting to notice that 
in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
 logical criterion also may be used in three senses -- of the
 agent, or the instrument, or the ``according to what''; the
 agent, for instance, may be a man, the instrument either sense
 perception or intelligence, and the ``according to what'' the
 application of the impression ``according to'' which the man
 proceeds to judge by means of one of the aforesaid
 instruments.''. The author than proceeds showing that
 even in this framework, knowledge still is non-apprehensible.
}.
Some relevant examples for our objective are the following.
\begin{itemize}
	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
		difference between knowledge and belief (as epistemological
		concepts), and the whole Doxastic logic defines in details how
		beliefs can be formalized.
	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
		of information and the difference with knowledge and belief.
	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
		tuple of assertions, beliefs, and facts.
\end{itemize}
For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
knowledge, beliefs, and the information or assertions it provides; where
knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
proposition known by an agent, such that: (i) knowledge requires belief, (ii)
knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
the only objective of information is to exchange beliefs between agents.  As
defined in \autocite{Hintikka1993Information}, ``information is specified by
specifying which alternatives concerning the reality it admits and which
alternatives excludes''. This means that if we consider a propositional
variable (which admits the two alternatives True/False) its information is
defined as believed to be True/False and not believed to be the opposite.  Due
to the definition of information and then with its relation to the
probabilistic correlation to truth/reality, we consider information in relation
with an agent's beliefs. Similarly, we consider beliefs to define the actual
behavior of an agent or a system; i.e. we consider behaviors as belief revision
systems.
The difference between knowledge and information as used by Hintikka, and facts
and assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
difficult to define
\autocite{Gettier2012knowledge}.
%\footnote{``\emph{Theaetetus}: [\ldots] He said
%that knowledge was true opinion accompanied by reason, but that unreasoning
%true opinion was outside of the sphere of knowledge; and matters of which there
%is not a rational explanation are unknowable -- yes, that is what he called
%them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
%[\ldots] the primary elements of which we and all else are composed admit of no
%rational explanation; for each alone by itself can only be named, and no
%qualification can be added, neither that it is nor that it is not, for that
%would at once be adding to it existence or non-existence, whereas we must add
%nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
%Theaetetus 201 \autocite{Plato1914Plato}}.
For example, Hintikka in
\autocite{Hintikka1993Information} states that ``A purely logical definition of
information is impossible''. However, a detailed formalization of those concepts is out
of the scope of this paper.

@@

\section{Prediction of Cybersecurity Weaknesses}\label{sec:theory}
Several standards mandate a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and
		``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
However, standards do not describe in detail how to perform a cybersecurity risk 
assessment and only vaguely define the overall objective, which
can be summarized as to provide an understanding of the potential cybersecurity risks.
Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculates the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
rely on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards require a cybersecurity risk assessment
which, in turn, mandates a metric to evaluate the cybersecurity risk. 
In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
and we identify a cybersecurity metric.

@@

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.8\textwidth]{mashyp.pdf}
	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
	\label{fig:mashyp}
\end{figure*}

>>>>

% Space hacks, ...yes, from outer space!
%\addtolength{\textfloatsep}{-6mm}
%\addtolength{\abovecaptionskip}{-3mm}
%\addtolength{\columnsep}{-1mm}
%\setlength{\evensidemargin}{0.1in}
%\setlength{\oddsidemargin}{0.1in}
%\setlength{\textwidth}{6.47in}
%\setlength{\textwidth}{5.1in}

@@

\author{
%	\IEEEauthorblockN{Marco Rocchetto}
%\IEEEauthorblockA{V-Research\\
%Verona, Italy\\
%Email: marco@v-research.it}
%\and
%\IEEEauthorblockN{Francesco Beltramini}
%\IEEEauthorblockA{V-Research\\
%Verona, Italy\\
%Email: francesco@v-research.it}
}

@@

There is no agreement between the research communities (e.g.
Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent.
However, the same ideas revolved around for thousands of years
\footnote{
It is interesting to notice that 
in \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
 logical criterion also may be used in three senses -- of the
 agent, or the instrument, or the ``according to what''; the
 agent, for instance, may be a man, the instrument either sense
 perception or intelligence, and the ``according to what'' the
 application of the impression ``according to'' which the man
 proceeds to judge by means of one of the aforesaid
 instruments.''. The author than proceeds showing that
 even in this framework, knowledge still is non-apprehensible.
}.
Some relevant examples for our objective are the following.
\begin{itemize}
	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
		difference between knowledge and belief (as epistemological
		concepts), and the whole Doxastic logic defines in details how
		beliefs can be formalized.
	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
		of information and the difference with knowledge and belief.
	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
		tuple of assertions, beliefs, and facts.
\end{itemize}
For our argument, as in \autocite{Santaca2016abf}, an agent is composed by its
knowledge, beliefs, and the information or assertions it provides; where
knowledge is defined, as in \autocite{Steup2020epistemology}, as a set of
proposition known by an agent, such that: (i) knowledge requires belief, (ii)
knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
the only objective of information is to exchange beliefs between agents.  As
defined in \autocite{Hintikka1993Information}, ``information is specified by
specifying which alternatives concerning the reality it admits and which
alternatives excludes''. This means that if we consider a propositional
variable (which admits the two alternatives True/False) its information is
defined as believed to be True/False and not believed to be the opposite.  Due
to the definition of information and then with its relation to the
probabilistic correlation to truth/reality, we consider information in relation
with an agent's beliefs. Similarly, we consider beliefs to define the actual
behavior of an agent or a system; i.e. we consider behaviors as belief revision
systems.
The difference between knowledge and information as used by Hintikka, and facts
and assertions as used in \autocite{Santaca2016abf}, are not precisely discussed
in \autocite{Santaca2016abf}.  Furthermore, those epistemological concepts are
difficult to define
\autocite{Gettier2012knowledge}.
%\footnote{``\emph{Theaetetus}: [\ldots] He said
%that knowledge was true opinion accompanied by reason, but that unreasoning
%true opinion was outside of the sphere of knowledge; and matters of which there
%is not a rational explanation are unknowable -- yes, that is what he called
%them -- and those of which there is are knowable. [\ldots] \emph{Socrates}:
%[\ldots] the primary elements of which we and all else are composed admit of no
%rational explanation; for each alone by itself can only be named, and no
%qualification can be added, neither that it is nor that it is not, for that
%would at once be adding to it existence or non-existence, whereas we must add
%nothing to it, if we are to speak of that itself alone.  [\ldots]'' Plato --
%Theaetetus 201 \autocite{Plato1914Plato}}.
For example, Hintikka in
\autocite{Hintikka1993Information} states that ``A purely logical definition of
information is impossible''. However, a detailed formalization of those concepts is out
of the scope of this paper.

@@

\section{Prediction of Cybersecurity Weaknesses}\label{sec:theory}
Several standards mandate a secure-by-design approach in which cybersecurity
shall be considered at the very early stages of the design process.  For
example,
\begin{enumerate}[noitemsep]
	\item DO-326A -- ``Airworthiness Security Process Specification''
		requires a cybersecurity risk assessment of the design and
		``are the only Acceptable Means of Compliance (AMC) by FAA
		\& EASA for aviation cyber-security airworthiness certification,
		as of 2019'' as pointed out by SAE in \autocite{SAE2019DO326A}.
	\item NIST 800-82 \autocite{Stouffer2011guide} -- ``guide to Industrial
		Control System (ICS) Security''
	\item J3061:2016-1 \autocite{SAE2016J3061} -- ``Cybersecurity Guidebook
		for Cyber-Physical Vehicle Systems'' defines ``set of
		high-level guiding principles for Cybersecurity as it relates
		to cyber-physical vehicle systems'' and states that
		``incorporate Cybersecurity into cyber-physical vehicle systems
		from concept phase through production, operation, service, and
		decommissioning''
\end{enumerate}
Standards do not describe in detail how to perform a cybersecurity risk 
assessment and only vaguely define the overall objective, which
can be summarized as to provide an understanding of the potential cybersecurity risks.
Roughly speaking, a cybersecurity risk assessment (e.g. CORAS \autocite{Lund2010model}) 
methodology starts from the (i) identification of assets, (ii) determines
the threat scenarios (correlating vulnerabilities, threats and incidents) e.g. 
threat diagrams in the CORAS
approach, (iii) calculates the risk as a function
of impact and likelihood of threat scenarios, and (iv) finally provides
treatments (mandating a partial re-design) and residual risks.
All the methodologies and tools we reviewed (e.g. Threatmodeler \autocite{Threatmodeler},
CORAS \autocite{Lund2010model}, SECRAM \autocite{De2015role})
rely on the expertise of the person who performs the risk assessment for
the identification of threats and for the quantitative estimation of risks.
As Herley puts it \autocite{Herley2016usenixvideo}, ``we don't have a test for security and
we don't have a metric for security'' while standards require a cybersecurity risk assessment
which, in turn, mandates a metric to evaluate the cybersecurity risk. 
In contrast, in this section, we define how to specify a CPS in the $\abftheory$-framework
and we identify a cybersecurity metric.

@@

\subsection{From Multi-Agent to Cyber-Physical Systems}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.7\textwidth]{mashyp.pdf}
	\caption{Mapping Epistemological Concepts to (Cyber-Physical) Systems Engineering}
	\label{fig:mashyp}
\end{figure*}
